---
title: "Assignment4"
author: "Robby Waxman"
date: "2/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning = F, results = "hide"}
#Install and import libraries if needed
if (!require('tidyverse')) install.packages('tidyverse')
library('tidyverse')
if (!require('devtools')) install.packages('devtools')
library('devtools')
if (!require('ggplot2')) install.packages('ggplot2')
library('ggplot2')
if (!require('ggmosaic')) install.packages('ggmosaic')
library('ggmosaic')
devtools::install_github("bcaffo/MRIcloudT1volumetrics")
MRIcloudT1volumetrics::readFileList("https://raw.githubusercontent.com/neuroconductor/kirby21.mricloud/master/inst/visit_1/127/kirby_3_1_ax_283Labels_M2_corrected_stats.txt")
```
## Question 1
Download the first Kirb21 subject data here. Use this package to read in the first subject. Calculate and report the intracranial volume: the sum of the level 1 type 1 volumes.

```{r}
fileLoc = system.file("extdata",
"kirby127a_3_1_ax_283Labels_M2_corrected_stats.txt",
package="MRIcloudT1volumetrics")
test = MRIcloudT1volumetrics::readSubject(fileLoc)
print(test[[1]])
print(sum(test[[1]]$volume))
```
So according to the calculation, there is a total volume in level 1 type 1 of 1378295


## Question 2
Read in the class data from here https://github.com/bcaffo/ds4bme/blob/master/data/classInterests.txt. Use ggplot to create bar plots of year and program.

First I do Year:
```{r}
filename <- "https://raw.githubusercontent.com/bcaffo/ds4ph-bme/master/data/classInterests.txt"
interests <- read.delim(filename, header = TRUE, sep = "\t")
ggplot(interests, aes(x=factor(Year)))+
  geom_bar(stat="count", width=0.7, fill="steelblue")+
  theme_minimal() + ggtitle("Bar Plot of Year") +
  xlab("Year") + ylab("Count")
```

Next I do Program
```{r}
ggplot(interests, aes(x=factor(Program)))+
  geom_bar(stat="count", width=0.7, fill="steelblue")+
  theme_minimal() + ggtitle("Bar Plot of Program") +
  xlab("Program") + ylab("Count")
```

## Question 3
Create a mosaic plot of the class data for year and program. It ends up a bit messy
```{r}
ggplot(data = interests) +
  geom_mosaic(aes(x = product(Year, Program), fill=Year)) + 
  labs(title='Mosaic Plot of Year and Program')
```

## Question 4
Download the gene expression dataset from here. Information about the data can be found in the sample info here. From the gene expression data matrix, subtract the rowmean from each row and the column mean from each column so that you have a new matrix that has been demeaned across both rows and columns.

```{r}
filename <- 'https://raw.githubusercontent.com/jhu-advdatasci/2018/master/data/GSE5859_exprs.csv'
dat <- read_csv(filename)
dat$X1 <- NULL
dat <- dat - rowMeans(dat)
dat <- dat - colMeans(dat)
#I decide to print out a few entries just to show the change
print(dat[1:4,1:4])
```

